---
title: "PCOS analysis"
author: "Harshita Singh"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggcorrplot)
```

##  Load the Data

```{r load-data}
pcos <- read_csv("PCOS_Lucas_Souza.csv")
head(pcos)
```

##  Data Overview

```{r summary}
glimpse(pcos)
summary(pcos)
colSums(is.na(pcos))
```

##  BMI by PCOS Diagnosis

```{r bmi-plot}
pcos$PCOS_Diagnosis <- as.factor(pcos$PCOS_Diagnosis)
ggplot(pcos, aes(x = PCOS_Diagnosis, y = BMI)) +
  geom_boxplot(fill = "skyblue") +
  labs(title = "BMI vs PCOS Diagnosis", x = "PCOS Diagnosis", y = "BMI")
```

##  Testosterone Levels by Diagnosis

```{r testosterone-plot}
ggplot(pcos, aes(x = PCOS_Diagnosis, y = `Testosterone_Level(ng/dL)`)) +
  geom_boxplot(fill = "salmon") +
  labs(title = "Testosterone Levels by PCOS Diagnosis", x = "PCOS Diagnosis", y = "Testosterone (ng/dL)")
```

##  Antral Follicle Count by Diagnosis

```{r follicle-plot}
ggplot(pcos, aes(x = PCOS_Diagnosis, y = Antral_Follicle_Count)) +
  geom_boxplot(fill = "lightgreen") +
  labs(title = "Antral Follicle Count by PCOS Diagnosis", x = "PCOS Diagnosis", y = "Follicle Count")
```

##  Menstrual Irregularity by PCOS Diagnosis

```{r period-plot}
ggplot(pcos, aes(x = Menstrual_Irregularity, fill = PCOS_Diagnosis)) +
  geom_bar(position = "dodge") +
  labs(title = "Menstrual Irregularity by PCOS Diagnosis", x = "Menstrual Irregularity", fill = "PCOS Diagnosis")
```

##  Correlation Heatmap

```{r correlation-plot}
pcos_numeric <- pcos %>%
  mutate(PCOS_Diagnosis = as.numeric(as.character(PCOS_Diagnosis))) %>%
  select(Age, BMI, `Testosterone_Level(ng/dL)`, Antral_Follicle_Count, Menstrual_Irregularity, PCOS_Diagnosis)

cor_matrix <- cor(pcos_numeric, use = "complete.obs")
ggcorrplot(cor_matrix, lab = TRUE, type = "lower")
```

## ✅  Conclusion

- Testosterone levels and follicle counts are higher in women diagnosed with PCOS.
- BMI and menstrual irregularities also show association with PCOS.
- Correlation analysis supports clinical insights about PCOS pathophysiology.

